---
title: "Parallelization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallelization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment  = "#>")
```

This tutorial explains different methods to run an analysis in parallel that are helpful when working on the cloud. All the scripts referenced below can be found in the [cloudDemo GitHub](https://github.com/LandSciTech/cloudDemo/tree/master/analyses/scripts)

## Running the analysis in parallel
So far our analysis is run sequentially so it will not be able to take advantage of the multiple cores available on a cloud machine. To make use of parallel computing we can set up our script to run each model on a separate core. Because this example is so simple I have added a delay in the function so we can see the benefit of running in parallel. There are many ways to do this and I will explain a few options. One option is to create a parallel backend in R using something like the `future` package. This can be quite straight forward and there are packages to connect this with familiar methods of iterating (eg for loops, `lapply` or `purrr::map`). This can get a bit tricky in that you are typically running only parts of the script on multiple cores and you need to make sure the right dependencies are available on the workers. `future` manages this for the most part but see https://future.futureverse.org/articles/future-4-issues.html for tips when that fails.  

### Using the future and furrr packages

`future` sets up the infrastructure needed to run things in the background. To initialize this you run `future::plan("multisession")` which will use all `future::availableCores()` by default. `furrr` is a parallelized version of `purrr` which allows you to iterate over elements in a list on separate cores. See script "analyses/02_run_model_furrr.R" for an example. 

### Using future and foreach

If you usually use for loops then the foreach package might be easier to use. This can be used with `future` to create the backend using the `doFuture` package. See script "analyses/scripts/03_run_model_foreach.R" for an example. 

## Running the tasks in parallel on multiple virtual machines

Instead of setting up a script to run in parallel on one machine we can split the job into multiple tasks that can run independently on separate machines. This is useful because each task will be contained in a separate process so if one iteration fails it will not impact the rest. The downside is that the whole script must be run in each task and each node will have to be set up with dependencies installed separately. 

To set this up we need to pass an argument to the R script that the task will run to tell it which iteration to run. The bash command to run one iteration would look like:

```
Rscript analyses/scripts/04_run_model_multitask.R 1
```
And in the R script "analyses/scripts/04_run_model_multitask.R" `commandArgs(trailingOnly = TRUE)` is used to access the number passed to the script (eg. 1). The number is then used to select the variable to test and to name the output of the task in a unique way. The output of each task is saved and downloaded so that we can compile the results and create our figure.

```{r}


r2_tab <- list.files(here::here("analyses/data/derived-data"), pattern = "*.csv", full.names = TRUE) |>
  purrr::map_dfr(read.csv) |> 
  dplyr::mutate(variable = reorder(variable, r.squared))

ggplot2::ggplot(r2_tab, ggplot2::aes(variable, r.squared))+
  ggplot2::geom_col()

```

For an example of how to run this whole analysis from the command line see
"analyses/scripts/run_azure_autoscale.sh" which also serves as an example of 
setting up autoscaling. Autoscaling automatically increases or decreases the 
number of nodes in a pool depending on how many tasks are waiting to start and 
how many are completed.  
